Case Study: 16S microbiome data (Schubert et al., 2014)

### 2018-02-26 datasets/ChIP-seq/ChIP-seq.Rmd
Make plot sizes and types uniform

```{r, chipseq-diag, fig.width=10, fig.height=3.5}
strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
```

@@

```{r, chipseq-plot, results="hide", width=15, height=15}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb, supplementary = FALSE)

rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)

>>>>

```{r, chipseq-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
```

```{r, chipseq-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
```

@@

```{r, chipseq-plot, results="hide", width=15, height=15}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb, supplementary = FALSE)
rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)

### 2018-02-26 datasets/GSEA/gsea-mouse.Rmd
Make plot sizes and types uniform

```{r, message=FALSE}
dsd_data <- "dsdHSCMPP.RData"
if(!file.exists(dsd_data)) {
    download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

@@

## run benchmark
sGSEAMouse <- buildBench(designBen, res, ftCols = "ind_covariate")
```

>>>>

```{r, message=FALSE}
dsd_data <- file.path(datdir, "dsdHSCMPP.RData")
if(!file.exists(dsd_data)) {
    download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

@@

## run benchmark
sb_file <- file.path(resdir, "sb-gsea-mouse.rda")
if (!file.exists(sb_file)) {
    sGSEAMouse <- buildBench(designBen, res, ftCols = "ind_covariate")
} else {
    sGSEAMouse <- readRDS(sb_file)
}
```

### 2018-02-26 datasets/microbiome/microbiome-baxter.Rmd
Make plot sizes and types uniform

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
out <- otu[rowSums(otu > 0) / nrow(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

@@

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)

@@

```{r, baxter-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.RData"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

@@

```{r, baxter-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, baxter-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

@@

```{r, baxter-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

@@

```{r, baxter-genus-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.RData"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = genus, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

@@

```{r, baxter-genus-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, baxter-genus-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

@@

```{r, baxter-genus-labelcovariate}
res <- res %>% mutate(ind_covariate = ubiquity)
```

@@

```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    print(nmids)
    bd <- initializeBenchDesign(nmids=nmids)
    ## remove ASH from comparison
    bd <- dropBMethod(bd, "ashs")

>>>>

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
otu <- otu[rowSums(otu > 0) / ncol(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

@@

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`).

@@

```{r, baxter-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.rda"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

@@

```{r, baxter-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, baxter-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, baxter-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, baxter-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

@@

```{r, baxter-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

@@

```{r, baxter-genus-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.rda"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = genus, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

@@

```{r, baxter-genus-ubi, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, baxter-genus-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, baxter-genus-abun, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, baxter-genus-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

@@

```{r, baxter-genus-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

@@

```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    ## use common BenchDesign w/out ASH
    bd <- initializeBenchDesign(nmids=nmids)
    bd <- dropBMethod(bd, "ashs")

### 2018-02-26 datasets/microbiome/microbiome-goodrich.Rmd
Make plot sizes and types uniform

```{r, goodrich-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=5, binwidth=0.05, numQ=4)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, goodrich-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

@@

```{r, goodrich-genus-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, goodrich-genus-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

>>>>

```{r, goodrich-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=5, binwidth=0.05, numQ=4)
```

```{r, goodrich-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, goodrich-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
```

```{r, goodrich-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

@@

```{r, goodrich-genus-ubi, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, goodrich-genus-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, goodrich-genus-abun, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, goodrich-genus-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

### 2018-02-26 datasets/microbiome/microbiome-schubert.Rmd
Make plot sizes and types uniform

```{r, schubert-ubi, fig.width=10, fig.height=3.5, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, schubert-abun, fig.width=10, fig.height=3.5, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

>>>>

```{r, schubert-ubi, fig.width=10, fig.height=3.2, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20, binwidth=0.05)
```

```{r, schubert-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

@@

```{r, schubert-abun, fig.width=10, fig.height=3.2, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, schubert-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```
